\vspace{-2cm} \LARGE Divergence-Free Smoothed Particle Hydrodynamics \\ \large Advanced Simulation and Visualization of Fluids in Computer Graphics

### 2016-12-14 report/paper.tex
Explained the simulation algorithm

\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
function performFluidSimulation
    for all particles $i$ do //Init neighbourhoods
        find neighbourhoods $N_i(0)$
    for all particles $i$ do //Init $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$
        compute densities $\rho_i(0)$
        compute factors $\alpha_i(0)$
    while (t < $t_{max}$) do //Start simulation loop
        adapt time step $\Delta{t}$ 
        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_g/m_i$
        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
        for all particles $i$ do //Update positions
            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
        for all particles $i$ do //Update neighbourhoods
            find neighbourhoods $N_i(t + \Delta{t})$
        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
            compute densities $\rho_i(0)$
            compute factors $\alpha_i(0)$
        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
\end{lstlisting}

@@

\subsection{Adapted time step}

@@

\subsection{Neighbourhood search}
    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
    Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
    The cutoff distance $h$ is the kernels smoothing radius.
    To make this faster a cell list was implemented.
    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $h$.
    Each cell spans on a surface in space.
    A particle belongs to a cell if it is inside the space the cell occupy.
    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see figure~\ref{fig:cellList}.

>>>>

Listing \ref{lst:simulation} drafts the produced simulation after following Bender et al.~\cite{bender} proposed method in detail. Upon starting the application will the particle neighbourhood $N_i$ be determined with a cell list solution explained in section \ref{section:neighbour}. For each particle will the alpha $\alpha_i$ and density $\rho_i$ parameters computed, both of which are commonly reused variables and thus worth storing to reduce computation cost. This is done before entering the simulation loop as initial values are required. During the loop will all these parameters be updated as the simulation progresses.

Once in the simulation at line $7$ will first the time step be determined through the CFL condition explained in section \ref{section:timestep}. The new velocities for each particle is then predicted through the included external forces which are propagated with ordinary euler integration.

\begin{lstlisting}[caption={Simulation algorithm},label={lst:simulation}, language=C++, mathescape=true, escapeinside={(*}{*)}] 
function performFluidSimulation
    for all particles $i$ do //Init neighbourhoods
        find neighbourhoods $N_i(0)$
    for all particles $i$ do //Init $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$
        compute densities $\rho_i(0)$
        compute factors $\alpha_i(0)$
    while (t < $t_{max}$) do //Start simulation loop
        adapt time step $\Delta{t}$ 
        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_g/m_i$
        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
        for all particles $i$ do //Update positions
            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
        for all particles $i$ do //Update neighbourhoods
            find neighbourhoods $N_i(t + \Delta{t})$
        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
            compute densities $\rho_i(0)$
            compute factors $\alpha_i(0)$
        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
\end{lstlisting}

The density solver uses the precomputed parameter $\alpha_i$ together with this prediction to compute the pressure forces in each neighbourhood such that it can correct the density error $\rho_i^* - \rho_i = 0$. This is further explained in \ref{densitySolver}.

After the particle position $\vect{x_i}$ has been updated are the neighbourhoods and the $\alpha_i$ and $\rho_i$ factors recalculated to reflect the new positions. Lastly, in line $19$ the divergence solver, much like the density solver, calculates the pressure forces to eliminate its error. In this case the divergence error $\frac{D\rho_i}{Dt}$ is corrected to be equal to $0$.

@@

\subsection{Adapted time step} \label{section:timestep}

@@

\subsection{Neighbourhood search} \label{section:neighbour}
    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
    Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
    The cutoff distance $h$ is the kernels smoothing radius.
    To make this faster a cell list was implemented.
    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $h$.
    Each cell spans on a surface in space.
    A particle belongs to a cell if it is inside the space the cell occupy.
    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see figure~\ref{fig:cellList}.
